Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-08-08 manuscript/coverletter/collabra_response.tex
work on response letter

\usepackage{longtable,booktabs,array}
\usepackage{float}
\begin{document}
	\name{Veronica Boyce}
\signature{\vspace{-35pt} Veronica Boyce, \\ on behalf of the author team}

@@

		I also address some of reviewer 1's specific repo-related concerns below; I suspect many of them would have been ameliorated had the reviewer been able to download the repo.

@@

		We have expanded the discussion section and re-organized it to have a more explicit discussion of limitations.

		Thank you for the reference, which we have incorporated, and we have included more information on the changes to the discussion below, in our response to reviewer 1.

		Also, see below for our changes and rebuttals to reviewer 1's thoughts.

@@

		%TODO discussion section
		We have expanded the discussion section to discuss the larger picture of different circumstances that motivate replications and rescues. We discuss the changes more below in response to reviewer 2's points.

@@

		We are sorry this error occurred, and that reviewer 1 didn't realize the files could be downloaded directly from the OSF copy (linked in the paper).  as the inability of reviewer 1 seems to have resulted in misunderstandings about the usability of files in the repo (which are much more understandable when opened on a computer).

		From googling, we believe this to be a windows issue, and to prevent this problem for future users we have added a note at the top of the README directing readers for how to address or sidestep the problem, namely:

		\revised{Note: If you are a windows user and get an error about overly long file names, please see [this stackoverflow post](https://stackoverflow.com/questions/22575662/filename-too-long-in-git-for-windows) or [this github issue](https://github.com/desktop/desktop/issues/17882) for how to fix it. Alternatively, you can download a zip of the repo from [the osf copy](https://osf.io/cyk5w/files/github) by clicking "Download this folder".}

@@

		See  https://github.com/langcog/251-rescue-projects/tree/main for the revised README.

@@

		As we now clarify in the updated README, these are archival copies of the reports (which were posted to Rpubs). From a cloned/downloaded copy of the repo, the html files can be opened in a browser to see the reports.

		% \revised{\begin{verbatim}
		 %		### First-replication-reports and Individual-rescue-reports
		 %		* archival copies of the reports from individual projects, in case the rpubs links break or are overwritten
		 %		* to view, download the repo and open the html files in a browser

		 %	\end{verbatim}}

@@

		The manuscript is written in an Rmd file that is reproducible. Each of the individual rescue projects was also written in a Rmd or qmd reproducible report (included in the individual project repos included in individual-rescue-projects/). We have now clarified this in the README.

	%	\revised{ \begin{verbatim}
	%		The overall manuscript can be reproduced by knitting manuscript/manuscript.Rmd
	%		* This file reads data from data/combined_data.csv and data/boyce_2023_data.csv
	%		* This file uses code/helper/parse_stats.R \end{verbatim}}

	%	\revised{\begin{verbatim}
	%			### First-replication-projects and Individual-rescue-projects
	%		* include the individual replication and rescue projects (aka github repos) respectively
	%		* the rescue projects repos have been checked and include data, code/reproducible report, and the materials to run the experiment\end{verbatim}}

@@

		%TODO bug students whose stuff wasn't all there

@@

		We agree that the frustration of original materials not being available or original instructions not being available is education and hopefully serves as an impetus for more sharing.

@@

		 %TODO extend the case study

		 We thank the reviewer for this idea. We have added brief mentions of the topics and outcomes of each of the studies. We disagree about the ordering, as some readers may be more interested in the bottom line overall results than the individual studies. To support the different tastes, we have added an overview of subsections at the beginning of results so that interested readers could skip to the casestudies if desired.

		 %TODO add mention of the casestudies existance

>>>>

\usepackage{longtable,booktabs,array}
\usepackage{float}

\providecommand{\tightlist}{%
	\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{hyperref}
\begin{document}
	\name{Veronica Boyce}
\signature{\vspace{-35pt} Veronica Boyce, \\ on behalf of the author team}

@@

		I also address reviewer 1's specific repository-related concerns more below; I suspect many of them would have been ameliorated had the reviewer been able to download the repository files.

@@

		We have added an explicit ``Limitations'' subsection to the discussion, which we discuss further below in our response to reviewer 1.

		Also, see below for our changes and rebuttals to reviewer 1's other points.

@@

		We have expanded the last part of the discussion section to acknowledge the much larger picture of different circumstances that motivate replications and rescues, and how these might be influenced by the prior beliefs of the investigators. We discuss the changes more below in response to reviewer 2's points.

@@

		We are sorry this error occurred, and that reviewer 1 didn't realize the files could be downloaded directly from the OSF copy (linked in the paper).  The inability of reviewer 1 to download and open the files seems to have resulted in misunderstandings about the usability of files in the repo (which are much more understandable when opened on a computer).

		From googling, we believe the long files names issue to be a windows issue, and to prevent this problem for future users we have added a note at the top of the README directing readers for how to address or sidestep the problem, namely:

		\revised{Note: If you are a windows user and get an error about overly long file names, please see [this stackoverflow post] (https://stackoverflow.com/questions/22575662/filename-too-long-in-git-for-windows) or [this github issue] (https://github.com/desktop/desktop/issues/17882) for how to fix it. Alternatively, you can download a zip of the repo from [the osf copy] (https://osf.io/cyk5w/files/github) by clicking "Download this folder".}

		Hopefully this warning will prevent future difficulties with accessing the repository.

@@

		See  https://github.com/langcog/251-rescue-projects/tree/main for the revised README, which I also quote here:

		\revised{
		\textbf{\large 251-rescue-projects}\label{rescue-projects}

			\textbf{Note:} If you are a windows user and get an error about overly
			long file names, please see
			\href{https://stackoverflow.com/questions/22575662/filename-too-long-in-git-for-windows}{this
				stackoverflow post} or
			\href{https://github.com/desktop/desktop/issues/17882}{this github
				issue} for how to fix it. Alternatively, you can download a zip of the
			repo from \href{https://osf.io/cyk5w/files/github}{the osf copy} by
			clicking ``Download this folder''.

			\textbf{\large Reproducing this project}\label{reproducing-this-project}

			The preregistration for this project is at https://osf.io/5qz7v

			The overall manuscript can be reproduced by knitting
			manuscript/manuscript.Rmd 
			\begin{itemize}
				\tightlist
				\item 
				This file reads data from
				data/combined\_data.csv and data/boyce\_2023\_data.csv
				\item  
				This file uses
				code/helper/parse\_stats.R
			\end{itemize}

			\textbf{data/combined\_data.csv has links to the write-ups,
				pre-registrations, and projects (code/data) for the replications and
				rescues.} Additionally, the project repos are contained in
			first-replication-projects and individual-rescue-projects. Archival
			copies of the write-ups are in first-replication-reports and
			individual-rescue-reports (download/clone the repo and then open the
			html files in a web browser to view them).

			The numeric results in data/combined\_data.csv (effect size measures,
			sample sizes, number of trials) were coded off of the papers and
			reports. Where the effect size measures were not presented in the needed
			format, recalculations were done (see reconstruct\_effect\_sizes for
			data and code for these).

			\textbf{\large Folder structure}\label{folder-structure}

			\textit{Code}\label{code}

			\begin{itemize}
				\tightlist
				\item
				initial\_criteria.R describes the process for determining which
				projects were eligible for rescue
				\item
				process.Rmd is a draft analysis script

				\begin{itemize}
					\tightlist
					\item
					it pulls the coded data from our google sheet and merges it with
					study-level variables from Boyce 2023 replication data to write
					data/combined\_data.csv
					\item
					it checks that we don't have parsing errors in our data format
					\item
					it does analyses (those in the manuscript have their analysis
					duplicated in manuscript/manuscript.Rmd)
				\end{itemize}
				\item
				helper/parse\_stats.R parses our raw effect size format and calculates
				standardized effect sizes and helper/stats.R wraps p\_original and
				prediction\_interval functions (used by process.Rmd but not
				manuscript.Rmd)
			\end{itemize}

			\textit{Data}\label{data}

			\begin{itemize}
				\tightlist
				\item
				combined\_data.csv is the raw data coding sheet with links to
				pre-registrations, etc
				\item
				parsed\_data.csv is a saved intermediate that has the results of
				parsing effect sizes and computing standardized effect sizes
				\item
				boyce\_2023\_data.csv is a copy of data from
				\href{https://royalsocietypublishing.org/doi/full/10.1098/rsos.231240}{Boyce
					et al 2023} which reports on the first replications
			\end{itemize}

			\textit{Manuscript}\label{manuscript}

			\begin{itemize}
				\tightlist
				\item
				manuscript.Rmd is the manuscript, with analysis code embedded
				\item
				manuscript.pdf is the result of knitting manuscript.Rmd
				\item
				251rescue.bib is the bibliography
				\item
				apa7.csl, generic\_article\_template.tex are helper files for
				formatting the manuscript
				\item
				coverletter/, title\_page.Rmd, and manuscript\_anon.Rmd are leftovers
				of the submission process
			\end{itemize}

			\textit{First-replication-projects and
				Individual-rescue-projects}\label{first-replication-projects-and-individual-rescue-projects}

			\begin{itemize}
				\tightlist
				\item
				include the individual replication and rescue projects (aka github
				repos) respectively
				\item
				the rescue projects repos have been checked and include data,
				code/reproducible report, and the materials to run the experiment
			\end{itemize}

			\textit{First-replication-reports and
				Individual-rescue-reports}\label{first-replication-reports-and-individual-rescue-reports}

			\begin{itemize}
				\tightlist
				\item
				archival copies of the reports from individual projects, in case the
				rpubs links break or are overwritten
				\item
				to view, download the repo and open the html files in a browser
			\end{itemize}

			\textit{reconstruct\_effect\_sizes}\label{reconstruct_effect_sizes}

			In a few cases, the effect size information we needed for comparable
			SMDs was not available from published papers or project write-ups, and
			we had to construct it. 
			\begin{itemize}
				\tightlist\item schectman\_es.Rmd calculates the effect size
			for the original
			\href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6634660/}{Schechtman
				et al.} The original paper reports a p-value but not exact values, but
			the authors were kind enough to send their matlab data files. This uses
			the three DataTest.mat files. Additionally, we looked at the
				rescue data to recalculate trial counts
				(subject\_data\_final\_full\_processed.csv) 
				\item 
			\href{https://www.pnas.org/doi/pdf/10.1073/pnas.2202657119}{Denis et al}
			is an additonal replication of Payne et al 2008; however the analysis
			reported in the paper is not the same as the analysis Payne et al used.
			Luckily data was available, so we calculate the corresponding result in
			denis\_es.Rmd using data in Expt1\_OverallRecognition.csv and
			Expt1\_SpecificRecognition.csv. 
			\item  For Hopkins et al, the main effect of
			interest is a coefficient in a linear model. However, coefficients
			cannot be converted to SMD easily, so we calculate the mean and sds
			instead in hopkins\_es.Rmd. The original data is in hopkins\_raw.csv,
			the replication and rescue data is sourced from their github repos.
		\end{itemize}

			\textit{So you want to rescue a project
				\ldots{}}\label{so-you-want-to-rescue-a-project}

			\begin{itemize}
				\item
				this is a copy of a list of potential reasons for replication failures
				shared with the rescuers in the early stages of the rescues
			\end{itemize}
		}

@@

		As we now clarify in the updated README (quoted above), these are archival copies of the reports (which were posted to Rpubs). From a cloned/downloaded copy of the repo, the html files can be opened in a browser to see the reports.

@@

		The manuscript is written in an Rmd file that is reproducible. Each of the individual rescue projects was also written in a Rmd or Qmd reproducible report (included in the individual project repos included in individual-rescue-projects/). We have now clarified this in the README.

@@

@@

		We agree that the frustration of original materials not being available or original instructions not being available is educational and hopefully serves as an impetus for more sharing.

@@

		 %TODO update after MS is potentially revised

		 We thank the reviewer for this idea. We have added an Appendix with brief descriptions of the topics and outcomes of each of the studies.

		  We disagree about the ordering, as some readers may be more interested in the bottom line overall results than the individual studies.

		 The contents of the Appendix is now as follows:

		 \revised{

		 	Write-ups of individual rescue and replication projects can be found in the repository at \url{https://osf.io/cyk5w/}.
		 	\newline

		 	Birch \& Bloom (2007) is about a false belief task comparing ignorance and plausible knowledge conditions.
		 	The first replication and an additional class replication both found results in the same direction as the original result, but much smaller.
		 	The rescue used a much larger sample size than both the original and the first replication and added an additional check to make sure the probabilities participants reported add up to 100\%.
		 	The rescue result was in the same direction as the original and prior replications, but did not find a significant effect.
		 	\newline

		 	Chou et al. (2016) studied whether a sense of control affects how much one feels physical pain. Participants were asked to recall experiences where they had either a low or high amount of control and then to report their level of physical pain. The first replication closely followed the methods of the original, but added a manipulation check borrowed from a different experiment in Chou et al. (2016). The manipulation check was successful, but there were no significant difference in pain level between the two groups. The rescue added pre-survey questions about pain levels to confirm that participants in the two groups had the same baseline understanding of pain. The rescue also did not find a significant difference in pain level between groups.
		 	\newline

		 	Craig \& Richeson (2014) had White Americans read an article about shifting racial demographics in the US, either with or without a framing that assuaged the potential threat, and found that reading the un-assuaged version led to the endorsement of more conservative positions.
		 	The first replication did not find a significant difference between conditions, and the rescuer speculated that this was due to a smaller sample size and a skew toward liberal political beliefs in the participant pool.
		 	The rescue recruited the same number of participants as the original study and dropped one of the conditions to increase power.
		 	Despite the increased power, the rescue did not replicate the original paper's finding.
		 	\newline

		 	Gong et al. (2019) looked at the interplay between age and willingness to donate in older adults in Hong Kong and found a three-way interaction between age, type of donation, and relationship to recipient.
		 	The first replication sampled US adults and did not find a significant interaction.
		 	The rescuer thought that cultural differences may have played a role and specifically sampled Asian American adults.
		 	The rescue did not recover the interaction.
		 	\newline

		 	Haimovitz \& Dweck (2016) manipulated parents' mindsets with ``failure-is-enhancing'' or ``failure-is-debilitating'' primes and looked at their reactions to a hypothetical situation in which a child received a failing grade.
		 	The first replication found a point estimate in the opposite direction to the original effect, but an additional class replication found a directionally consistent effect.
		 	The rescue added manipulation checks which were not present in the original study or either replications.
		 	The manipulation checks passed, but the rescue study did not recover the target effect.
		 	\newline

		 	Hopkins et al. (2016) examined whether people prefer reductive explanations, which explain a result in one science with a more ``fundamental'' science, to horizontal explanations, which explain a result in one science in terms of the same science.
		 	While the original study found a preference for reductive explanations, the first replication did not.
		 	The replication's effective sample size was smaller than the original because a large proportion of participants failed the attention check.
		 	The rescue increased made the attention check slightly easier and recruited a larger sample to account for failures, but still did not replicate the original effect.
		 	\newline

		 	Jara-Ettinger et al. (2022) looked at shape bias in US adults across 3 shape exemplars and found a preference to extend nouns on the basis of shape rather than color or material.
		 	The replication found an above-chance preference for shape, but with very high item-level variability.
		 	The rescuer found that the first replication used stylized versions of the original stimuli taken from a different part of the paper.
		 	The rescue used the same images of stimuli as the original, but still found strong item-level effects and relative preferences for color and material.
		 	\newline

		 	Paxton et al. (2012) found that participants who completed the Cognitive Reflection Test (CRT) before the moral judgement task gave more utilitarian answers than those who did the moral judgment task before the CRT.
		 	The replication rewrote some of the CRT questions, adding a sub-condition where some of the questions featured characters with specific names.
		 	The first replication failed to find an effect.
		 	The rescuer removed the sub-condition and used questions from the CRT-2, out of concern that Prolific participants would be familiar with the original CRT questions.
		 	The rescue also failed to find the effect.
		 	\newline

		 	Porter et al. (2016) looked at how linguistic intergroup bias influences judgments of shared group membership.
		 	The first replication found an effect directionally inconsistent with the original result.
		 	The rescuer obtained the direct wordings of the questions from the authors of the original paper, and noticed that they were substantially different from questions in the first replication.
		 	The rescue used the same wordings as the original paper and replicated the effect, although the manipulation check failed.
		 	\newline

		 	Schechtman et al. (2010) looked at how widely people generalize tones that previously have been paired with losses or gains, and found that the loss-paired tones were generalized more.
		 	The first replication found a statistically significant effect in the opposite direction of the original study.
		 	It also found that people took much longer to learn which tones were associated with gains vs.~losses compared to the original.
		 	The rescuer made several changes, including asking questions about the sound quality of the participant's computer.
		 	The original study was conducted in person while the replication and rescue were online, so differences in speaker and headphone quality might have played a role in the different result.
		 	The rescue found an effect in the same direction as the 1st replication (contrary to the original study).
		 	\newline

		 	As discussed in methods, Tarampi et al. (2016) had participants do a timed navigation task and found an interaction between participant gender and the framing of the task (social or spatial) on performance.
		 	The result from the first replication was both statistically insignificant and directionally inconsistent.
		 	The original study was conducted in person, while the replication was conducted online, which might have contributed to the inconsistency.
		 	The rescue was also conducted online, but the rescuer made a series of user interface changes, a comprehension check, and exclusion criteria to ensure that participants understood the instructions, found the task convenient to interact with, and were engaged throughout the experiment.
		 	Still, the rescue did not recover the interaction.
		 	\newline

		 	Todd et al. (2016) looked at participants speed at classifying toys and guns as a function of whether they had just need a White or Black boy's face.
		 	Todd et al. (2016) found an interaction where Black faces primed faster ``gun'' responses and White faces faster ``toy'' responses.
		 	When we reached out to the original authors, they offered to run a parallel rescue on their student subject pool, which we have here as an additional replication.
		 	The first replication did not find the interaction, but both the better-powered rescue and the original authors' replication did.
		 	\newline

		 	Yeshurun \& Levy (2003) looked at the relationship between spatial attention and temporal resolution.
		 	They measured the threshold where two flashes of light are perceived as one flash versus two as a function of whether the location of the flashes was pre-cued or not.
		 	Yeshurun \& Levy (2003) found that cueing led to less temporal resolution.
		 	This was independently replicated by Chica \& Christie (2009), but was not replicated by either the student replication or the rescue.
		 	Both the replication and rescue used in-person participants because the study relies on precise timing and details of the screen which are difficult to control online.
		 	The first replication had a smaller sample size than the original, while the rescue matched the original study's sample size.

		 }

### 2024-08-08 manuscript/manuscript.Rmd
work on response letter

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (<https://osf.io/dfwk2>).
In general, we encouraged students to add manipulation checks as appropriate, and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of @paxton2012 switched from using the Cognitive Teflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2 which is less overused but attempts to measure the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants click a drop down menu to select left or right for each turn.
The rescue removed the unwieldy drop down menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

>>>>

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (<https://osf.io/dfwk2>).
In general, we encouraged students to add manipulation checks as appropriate, and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of @paxton2012 switched from using the Cognitive Reflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2 which is less overused but attempts to measure the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants click a drop down menu to select left or right for each turn.
The rescue removed the unwieldy drop down menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

### 2024-08-08 manuscript/manuscript.tex
work on response letter

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (\url{https://osf.io/dfwk2}).
In general, we encouraged students to add manipulation checks as appropriate, and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of Paxton et al. (\citeproc{ref-paxton2012}{2012}) switched from using the Cognitive Teflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2 which is less overused but attempts to measure the same construct (\citeproc{ref-thomson2016}{Thomson \& Oppenheimer, 2016}).
The rescue of Jara-Ettinger et al. (\citeproc{ref-jara-ettinger2022}{2022}) discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
Tarampi et al. (\citeproc{ref-tarampi2016}{2016}) had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants click a drop down menu to select left or right for each turn.
The rescue removed the unwieldy drop down menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

>>>>

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (\url{https://osf.io/dfwk2}).
In general, we encouraged students to add manipulation checks as appropriate, and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of Paxton et al. (\citeproc{ref-paxton2012}{2012}) switched from using the Cognitive Reflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2 which is less overused but attempts to measure the same construct (\citeproc{ref-thomson2016}{Thomson \& Oppenheimer, 2016}).
The rescue of Jara-Ettinger et al. (\citeproc{ref-jara-ettinger2022}{2022}) discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
Tarampi et al. (\citeproc{ref-tarampi2016}{2016}) had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants click a drop down menu to select left or right for each turn.
The rescue removed the unwieldy drop down menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.
